---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# cards

<!-- badges: start -->
[![R-CMD-check](https://github.com/jabenninghoff/cards/workflows/R-CMD-check/badge.svg)](https://github.com/jabenninghoff/cards/actions)
[![lint](https://github.com/jabenninghoff/cards/workflows/lint/badge.svg)](https://github.com/jabenninghoff/cards/actions)
[![Codecov test coverage](https://codecov.io/gh/jabenninghoff/cards/branch/main/graph/badge.svg)](https://app.codecov.io/gh/jabenninghoff/cards?branch=main)
<!-- badges: end -->

## Overview

**`r desc::desc_get_field("Title")`:** `r desc::desc_get_field("Description")`

Feel free to use and/or fork this project!

## Installation

You can install the development version of cards from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("jabenninghoff/cards")
```

Or using renv:

``` r
# install.packages("renv")
renv::install("jabenninghoff/cards")
```

## Development

- Changelog: See "Changelog" or `NEWS.md`.
- Planned: [TODO](TODO.md)

## Usage

The cards package supports random draws and evaluation of poker rank [category](https://en.wikipedia.org/wiki/List_of_poker_hands):

```{r example}
library(cards)

deck <- new_deck_df()
replicate(50, {
  hand <- deal_hand_df(deck)
  paste0(print_hand(hand), ": ", eval_hand(hand))
})
```
