---
title: "Benchmarks"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Benchmarks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Measure the performance of different implementations of `cards` using `bench::mark()`.

```{r setup, message = FALSE, warning = FALSE}
library(cards)
```

# Data Frame

Benchmark the initial implementation using `data.frame` compared to an `integer()` approach similar to [PH Evaluator](https://github.com/HenryRLee/PokerHandEvaluator) [`card.py`](https://github.com/HenryRLee/PokerHandEvaluator/blob/master/python/phevaluator/card.py).

## New Deck

Create a new deck using `new_deck_df()` and an integer vector.

```{r new_deck_df}
deck <- new_deck_df()
deck_int <- 0:51
bench::mark(new_deck_df())
bench::mark(0:51)
```

While `new_deck_df()` is not designed to be called frequently, using an integer vector is much faster.

## Deal

Compare performance of `deal_hand_df()` to sampling integers:

```{r deal_hand_df}
bench::mark(deal_hand_df(deck))
bench::mark(sample(deck_int, 5))
```

`deal_hand_df()` is about 7 times slower than `sample()`.

## Print

Test performance of `print_hand_df()` against a simple function that prints cards based on integers.

```{r print_hand_df}
test_hand <- deal_hand_df(deck)
bench::mark(print_hand_df(test_hand))

print_hand_int <- function(h) {
  cards <- paste0(rep(c(2:9, "T", "J", "Q", "K", "A"), each = 4), c("C", "D", "H", "S"))
  paste0(cards[h + 1], collapse = " ")
}
test_hand_int <- sample(0:51, 5)
bench::mark(print_hand_int(test_hand_int))
```

`print_hand_df()` is 14-15 times slower than the integer approach.

## Evaluate

Test performance of `eval_hand_df()` with a single hand and with randomly selected hands:

```{r eval_hand_df}
bench::mark(eval_hand_df(test_hand))
bench::mark(eval_hand_df(deal_hand_df(deck)))
```

As expected for a naive poker hand evaluator, performance of `eval_hand_df()` is poor compared to fast algorithms.

## Summary

An implementation using integer would likely be much faster than the first implementation using `data.frame`. Rank and suit can be derived using integer division and modulo arithmetic respectively, and `tabulate()` is a faster replacement for `rle()`.

```{r rle_tabulate}
0:51 %/% 4
tabulate(0:51 %/% 4 + 1, 13)
0:51 %% 4
tabulate(0:51 %% 4 + 1, 4)

bench::mark(rle(sort(sample(0:51, 5) %/% 4 + 1)))
bench::mark(tabulate(sample(0:51, 5) %/% 4 + 1, 13))
```

Note that the tabulate approach is 7 times faster than sorting and run length encoding.
