---
title: "Video Poker"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Video Poker}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Simulate "Full Pay Jacks or Better" [optimal strategy](https://wizardofodds.com/games/video-poker/strategy/jacks-or-better/9-6/optimal/) video poker. As the linked article notes:

> The following strategy is for full pay Jacks or Better video poker. "Full Pay" designates the following paytable, per coin based on five coins bet, which returns 99.54% of money bet assuming optimal strategy.

```{r setup, message = FALSE, warning = FALSE}
library(cards)
library(tibble)
library(dplyr)

deck <- new_deck()
runs <- 2e6
```

# One Hand

Run simple simulations with a single hand, without drawing new cards.

```{r hands}
hands <- replicate(runs, eval_hand(deal_hand(deck)))
```

Calculate payouts, using the table from the [optimal strategy article](https://wizardofodds.com/games/video-poker/strategy/jacks-or-better/9-6/optimal/) and these [pay tables](https://web.archive.org/web/20221125191618/https://www.gamblingsites.org/casino/video-poker/pay-tables/):

Hand             Payout
--------------- -------
Royal Flush     800
Straight Flush  50
Four of a Kind  25
Full House      9
Flush           6
Straight        4
Three of a Kind 3
Two Pair        2
Jacks or Better 1

Payout is per coin when betting 5 coins.

```{r payout}
payout <- hands |>
  as_tibble_col(column_name = "hand") |>
  mutate(payout = case_match(
    hand,
    c("high_card", "one_pair") ~ 0,
    "jacks_better" ~ 1,
    "two_pair" ~ 2,
    "three_ofakind" ~ 3,
    "straight" ~ 4,
    "flush" ~ 6,
    "full_house" ~ 9,
    "four_ofakind" ~ 25,
    "straight_flush" ~ 50,
    "royal_flush" ~ 800
  ))
```

Filter only Royal Flushes:

```{r royal_flush}
payout |>
  filter(payout == 800)
```

Calculate the total payout, and percentage return:

```{r}
total <- payout |>
  pull(payout) |>
  sum()

total
total / runs
```

A simple "one hand" simulation performs much poorer than the expected return of 99.54% for optimal strategy.
